% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%

\usepackage{xcolor}
\usepackage{caption}
%\newcommand{\edits}[1]{\textcolor{red}{#1}}
%\newcommand\boldblue[1]{\textcolor{blue}{\textbf{#1}}}
\newcommand\hn[1]{\textcolor{cyan}{#1}}
\newcommand\todo[1]{\textcolor{red}{#1}}
\newcommand\newtext[1]{\textcolor{violet}{#1}}
\newcommand\notes[1]{\textcolor{green}{#1}}
\setcounter{secnumdepth}{3}
%% ceurart loads most packages. Add only what is genuinely needed.
\usepackage{booktabs}        % \toprule, \midrule, \bottomrule
\usepackage{listings}        % code listings (SPARQL, Turtle)
\usepackage{microtype}       % better justification
\usepackage{amssymb}         % for checkmarks

\begin{document}
%
\title{{\texttt{GeMeA}: Knowledge Graph Testbed for 27M+ German Cultural Heritage Objects}}
%
\titlerunning{GEMEA}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mary Ann Tan\inst{1,2}\orcidID{0000-0003-3634-3550} \and
Etienne Posthumus\inst{1,3}\orcidID{0000-0002-0006-7542} \and
Genet Asefa Gesese\inst{1,2}\orcidID{0000-0003-3807-7145} \and
Harald Sack\inst{1,2}\orcidID{0000-0001-7069-9804}}
%
\authorrunning{M. Tan et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{FIZ Karlsruhe – Leibniz Institute for Information Infrastructure, Germany 
\email{\{ann.tan|genet-asefa.gesese|harald.sack\}@fiz-karlsruhe.de} \and
Karlsruhe Institute of Technology, Institute AIFB, Germany \and
\email{ep@poz.org}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{00-abstract}
\keywords{knowledge graphs \and Linked Open Data \and digital humanities \and ontology alignment \and entity linking \and research infrastructure \and Model Context Protocol.}
\end{abstract}
%
%
%
\input{10-intro}
\input{20-related}
\input{30-resource}
\input{40-quality}
\input{50-usage}
\input{60-impact}
\input{70-sustainability}
\input{80-conclusion}

\subsubsection*{Resource Availability Statement.}
\todo{Persistent URI, SPARQL endpoint URL, SHMARQL browser URL, source code repository, CC BY 4.0 licence statement.}

\begin{credits}
\subsubsection{\ackname} The authors thank Christian Nennemann for developing the QLever MCP server. \todo{Funding acknowledgments to be added.}

\subsubsection{\discintname}
The authors have no competing interests to declare that are relevant to the content of this article.

\subsubsection*{Declaration on Generative AI.}
\todo{AI use statement to be added.}

\subsubsection*{CRediT.}
\textbf{Mary Ann Tan}: Conceptualization, Data curation, Formal analysis, Investigation, Methodology, Resources, Software, Validation, Visualization, Writing -- original draft, Writing -- review \& editing.
\textbf{Etienne Posthumus}: Software, Writing -- review \& editing.
\textbf{Genet Asefa Gesese}: Writing -- review \& editing.
\textbf{Harald Sack}: Project administration, Supervision, Writing -- review \& editing.
\end{credits}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{99-bibliography}

\end{document}
