%% 30-resource.tex — Resource Description

\section{Resource Description}
\label{sec:resource}

GEMEA is a knowledge graph over 26 million objects from the German Digital Library (DDB), organized into named graphs by provider institution and accessible via a public SPARQL endpoint, a Linked Data browser, and an MCP/MCPO agentic interface.
The resource is built in four layers: (1) an EDM knowledge graph loaded into QLever, (2) a \textbf{MOCHO}-based ontology alignment layer mapping EDM to RDA/FRBR on a representative subset, (3) a SHMARQL Linked Data browsing interface, and (4) an MCP interface on the QLever endpoint.

\subsection{EDM Knowledge Graph}
\label{sec:resource:kg}

The source data is the DDB corpus: JSON-LD exports of EDM records aggregated from \todo{N} German cultural heritage institutions across five sectors --- Bibliothek (library), Archiv (archive), Museum, Mediathek (media library), and Denkmal (monument preservation).
Records are converted from DDB JSON-LD to N-Triples using rdf2jsonld~\cite{rdf2jsonld}, which normalizes EDM predicates, repairs malformed URIs, and partitions output by provider institution.

\paragraph{Named graph structure.}
The QLever index is organized as named graphs following the scheme \texttt{http://gemea.ddb.de/graph/\{provider\_id\}}, one per provider institution.
A summary graph at \texttt{http://gemea.ddb.de/graph/meta} holds VoID and DCAT dataset descriptors.
Named graph partitioning preserves provenance at the provider level and enables per-institution queries without full-corpus scans.

\paragraph{Scale.}
\todo{Run \texttt{SELECT (COUNT(*) AS ?n) WHERE \{ ?s ?p ?o \}} per named graph and fill in: total triple count, ProvidedCHO count, provider count, named graph count.}

\paragraph{Object types and sectors.}
\todo{Insert sector and mediatype distribution table from QLever once stats are available.}

\paragraph{VoID descriptor.}
\todo{Insert VoID snippet (void:Dataset with triples, distinctSubjects, properties, sparqlEndpoint, dataDump).}

\subsection{Mocho Ontology Alignment PoC}
\label{sec:resource:mocho}

EDM is designed for cross-institutional aggregation rather than for KG reasoning or bibliographic work-level grouping.
\textbf{MOCHO}~\cite{mocho} maps EDM to RDA/FRBR-compatible representations: \texttt{edm:ProvidedCHO} is typed as \texttt{frbr:Manifestation}; agents are split into \texttt{frbr:Person} and \texttt{frbr:CorporateBody} by GND type; EDM predicates are mapped to RDA vocabulary (e.g., \texttt{dc:title} $\to$ \texttt{rda:title}, \texttt{dc:creator} $\to$ \texttt{rda:creator}).
Where a GND Werk URI is available (from \texttt{link\_gnd\_works.py}), \textbf{MOCHO} groups co-manifesting ProvidedCHOs into a \texttt{mocho:Work} entity.

The alignment PoC is run on a representative subset of the DDB corpus.
\todo{Subset size (CHO count), classes produced, triple count, mapping quality summary (direct vs.\ approximate matches, known scope boundaries).}

Real EDM encodings surface scope boundaries in the initial alignment rules --- most notably in \texttt{dc:type} values, which carry locally-defined free-text strings in several sectors rather than the controlled vocabulary terms the rules assumed.
These boundaries are documented in Section~\ref{sec:quality}.

\paragraph{Corpus-driven dispatch design.}
A foundational constraint of the alignment methodology is that EDM property semantics are sector-dependent: \texttt{dc:type} carries controlled-vocabulary terms in library records, free-text strings in museum records, and structural hierarchy identifiers in archive records.
These differences are not declared in the EDM schema specification and are invisible to schema-level or structure-based alignment tools.
Reliable automated dispatch therefore requires corpus-based analysis at two levels: coarse-grained stratification by sector~$\times$~mediatype to determine which classification signals are available and reliable per stratum, and fine-grained value distribution analysis to discover the actual semantics in use within each stratum.
Concretely, before dispatch logic is implemented, classification signal coverage is measured per sector~$\times$~mediatype stratum across the Goethe-Faust corpus.
Two signals are available per record: \texttt{ddb:hierarchyType} (htype) and \texttt{dc:type}.
Coverage varies substantially by stratum: Archive (sparte001) text records carry both signals in 99.6\% of cases; Library (sparte002) text records are htype-only in 31\%; Museum, Media Library, and Monument sectors (sparte003--006) carry no htype at all.
This coverage profile determines dispatch priority per stratum --- htype-first for Archive and Library structural classes, dc:type-first for the remaining sectors --- rather than applying a uniform fallback chain across all records.
The full coverage table is archived in \texttt{output/dispatch\_signal\_ratio.csv} (script: \texttt{scripts/dispatch\_signal\_ratio.py}).

A further constraint shaped the alignment methodology: the target ontology was not fixed before transformation began.
New \textbf{MOCHO} classes --- \texttt{mocho:ImageWork}, \texttt{mocho:ImmovableWork}, \texttt{mocho:ImageObject} --- were introduced in direct response to patterns surfaced during corpus analysis, making the workflow iterative: corpus inspection identified a gap, the ontology was extended, alignment rules were updated.
This is characteristic of large-scale cultural heritage transformation, where the correct ontological structure for heterogeneous institutional data is itself a research question that cannot be settled in advance.
Automated schema-matching tools presuppose a stable target ontology and cannot signal when an ontology gap, rather than a nearest-available substitute, is the correct resolution; manual curation was therefore not a substitute for automation but a methodological requirement.

\subsection{SHMARQL Linked Data Interface}
\label{sec:resource:shmarql}

SHMARQL~\cite{shmarql} is a lightweight Linked Data publishing layer deployed as a Docker container (\texttt{ghcr.io/epoz/shmarql}).
Entity URIs are dereferenceable, RDF is served with content negotiation (HTML, Turtle, JSON-LD), and a built-in SPARQL query editor is included.
No custom frontend development is required.

For GEMEA, SHMARQL is deployed as a proxy over the QLever endpoint.
The deployment pattern was validated on the Goethe-Faust corpus~\cite{goethefaust} before being adapted for GEMEA's larger index.
The GEMEA SHMARQL instance is available at \todo{[SHMARQL URL]}.

\todo{Insert 1--2 screenshots of the SHMARQL interface: entity page and SPARQL editor.}

\subsection{MCP/MCPO Agentic Access Layer}
\label{sec:resource:mcp}

The Model Context Protocol (MCP)~\cite{mcp2024} is an open standard for exposing tools and data sources to AI agents.
GEMEA's QLever SPARQL endpoint is registered as an MCP server using \texttt{mcp-server-qlever}~\cite{mcpqlever}, making the KG queryable by any MCP-compatible agent or LLM client without custom integration code.
An MCPO~\cite{mcpo} wrapper further exposes the endpoint as an OpenAPI-compatible HTTP interface for frameworks that do not natively support MCP.

The registration is reproducible via \texttt{./setup.sh mcp-add} in the GEMEA Docker Compose stack, following the pattern established in the Goethe-Faust deployment.
\todo{Inventory of MCP tools exposed by mcp-server-qlever: list tool names and descriptions.}

\paragraph{Example agent interaction.}
\todo{Insert 2 example agent interactions (MCP tool calls + responses) once mcp-add is run.}
