\section{Quality and Validation}
\label{sec:quality}

% TODO (~1–2 pages)
% - Completeness: coverage of DDB corpus (% of total objects loaded)
% - URI well-formedness: results from rdf2jsonld URI validator (repair rate, drop rate)
% - mocho mapping quality: direct vs. approximate matches; coverage per ontology
% - mocho was built with ODK and reasoned with HermiT~\cite{motik2009hypertableau}
% - SHACL constraints / SPARQL ASK validation queries
% - Consistency checks: duplicate detection, dangling references
% - Known limitations: unresolved entities, sparse geographic data, temporal gaps
% - Metrics table
\textbf{MOCHO} was built with the Ontology Development Kit (ODK).
The ontology requires OWL~2~DL expressivity: it makes use of universal restrictions
(\texttt{owl:allValuesFrom}), cardinality constraints, complement and disjointness axioms,
and property chains, which collectively place it outside the OWL~2~EL profile and preclude
the use of consequence-based reasoners such as ELK~\cite{kazakov2014elk}.

In \textbf{MOCHO}~1.0, classification was performed with
HermiT~\cite{motik2009hypertableau}, whose hypertableau algorithm provides complete
OWL~2~DL reasoning but proved slow for our ontology.
Subsequent versions use \textbf{Openllet}~\cite{sirin2007pellet}\footnote{\url{https://github.com/Galigator/openllet}}, the actively maintained
open-source fork of Pellet, for the following reasons:
%
\begin{itemize}
  \item \textbf{ROBOT integration.} Openllet is supported natively by
        \texttt{robot reason -{}-reasoner Openllet}, fitting directly into the ODK
        build pipeline without additional server infrastructure.
  \item \textbf{Explanation support.} Openllet can produce justifications for
        entailments and inconsistencies, which is essential for debugging alignment
        axioms across the RDA, EDM, and CIDOC-CRM vocabularies imported by \textbf{MOCHO}.
  \item \textbf{Active maintenance.} Konclude~\cite{steigmiller2014konclude} offers
        faster parallel tableau classification but has not been updated since 2020;
        Openllet receives ongoing releases. At \textbf{MOCHO}'s current scale
        (334 classes, 456 object properties) the parallelism advantage of Konclude
        does not outweigh the tooling and maintenance cost.
\end{itemize}
%
Materialised inferences are exported as RDF triples and loaded into QLever alongside
the DDB corpus; no runtime reasoning is performed by the triplestore.

\todo{Quality and Validation.}
