%% 50-usage.tex — Usage and Reusability

\section{Usage and Reusability}
\label{sec:usage}

\subsection{SPARQL Access}

A public read-only SPARQL 1.1 endpoint is available at \todo{[endpoint URL]}, backed by QLever.
QLever's integrated text and SPARQL support allows queries that combine keyword matching with graph constraints in a single request via \texttt{ql:contains-entity}.

\paragraph{Query 1: Objects associated with a given agent (GND URI).}
\begin{lstlisting}[language=SPARQL, basicstyle=\ttfamily\small, breaklines=true]
PREFIX edm: <http://www.europeana.eu/schemas/edm/>
PREFIX dc:  <http://purl.org/dc/elements/1.1/>
SELECT ?cho ?title WHERE {
  ?agg edm:aggregatedCHO ?cho .
  ?cho dc:creator <http://d-nb.info/gnd/118540238> ;
       dc:title   ?title .
}
LIMIT 100
\end{lstlisting}

\paragraph{Query 2: Objects created in a given place and decade.}
\begin{lstlisting}[language=SPARQL, basicstyle=\ttfamily\small, breaklines=true]
PREFIX edm:     <http://www.europeana.eu/schemas/edm/>
PREFIX dc:      <http://purl.org/dc/elements/1.1/>
PREFIX dcterms: <http://purl.org/dc/terms/>
SELECT ?cho ?title ?date WHERE {
  ?cho edm:currentLocation <http://sws.geonames.org/2925533/> ;
       dc:title             ?title ;
       dcterms:issued       ?date .
  FILTER(?date >= "1800"^^xsd:gYear && ?date < "1810"^^xsd:gYear)
}
LIMIT 100
\end{lstlisting}

\paragraph{Query 3: Object count per provider institution.}
\begin{lstlisting}[language=SPARQL, basicstyle=\ttfamily\small, breaklines=true]
PREFIX edm: <http://www.europeana.eu/schemas/edm/>
SELECT ?provider (COUNT(?cho) AS ?count) WHERE {
  ?agg edm:dataProvider ?provider ;
       edm:aggregatedCHO ?cho .
}
GROUP BY ?provider ORDER BY DESC(?count)
\end{lstlisting}

\paragraph{Query 4: One-hop neighborhood of a ProvidedCHO.}
\begin{lstlisting}[language=SPARQL, basicstyle=\ttfamily\small, breaklines=true]
PREFIX edm: <http://www.europeana.eu/schemas/edm/>
SELECT ?p ?o WHERE {
  { <http://www.deutsche-digitale-bibliothek.de/item/EXAMPLE123>
    ?p ?o . }
  UNION
  { ?o ?p
    <http://www.deutsche-digitale-bibliothek.de/item/EXAMPLE123> . }
}
\end{lstlisting}

\subsection{Linked Data Browser}

The SHMARQL interface (Section~\ref{sec:resource:shmarql}) provides point-and-click navigation over GeMeA without requiring SPARQL knowledge.
Entity URIs for ProvidedCHOs, agents, places, and time spans are dereferenceable: a browser request renders an HTML entity page with outgoing and incoming links; HTTP content negotiation returns Turtle or JSON-LD for machine clients.

\subsection{Research Applications}

GeMeA supports the following LLM + KG research use cases.

\paragraph{KG-grounded retrieval.}
The SPARQL endpoint and MCP interface support evaluation of KG-grounded RAG pipelines, entity-linked question answering, and hybrid text + structured query benchmarks over a real-world CH corpus at scale.
QLever's integrated text search (\texttt{ql:contains-entity}) enables single-query evaluation of text and graph constraints without a separate full-text search layer.

\paragraph{Entity embedding and link prediction.}
The 26M entity graph --- agents, places, timespans, providers, and \texttt{mocho:Work} groupings --- provides a benchmark for entity embedding methods and link prediction.
Named graph structure supports provenance-aware experiments that distinguish source triples from enrichment triples.

\paragraph{Explainability.}
Provenance-tracked named graphs and PROV-O\todo{cite here!} traces on GND-linked triples support evaluation of KG-grounded explanation methods, including tracing a retrieval result back to its source institution, enrichment pathway, and match confidence.

\paragraph{Agentic KG access (\texttt{framework.trails}).}
The MCP/MCPO interface supports development and evaluation of multi-step KG reasoning agents.
An agent with access to the GeMeA MCP server can resolve entity queries, explore graph neighborhoods, and compose multi-hop SPARQL queries without manual endpoint configuration.
\todo{Insert 2 example MCP agent interactions.}

\paragraph{Ontology evaluation (\texttt{onto-eval}).}
The \textbf{MOCHO}-aligned subgraph --- with documented alignment coverage and scope boundaries --- provides a grounded evaluation target for ontology evaluation methods and a benchmark for alignment improvement.

\subsection{Self-Hosting}

The full GeMeA stack is deployable via Docker Compose.
\todo{Hardware requirements, stack services (QLever, SHMARQL, MCP proxy), and link to deployment documentation.}
