%% 70-sustainability.tex — Sustainability

\section{Sustainability}
\label{sec:sustainability}

\paragraph{Versioning.}
Per-item provenance is recorded at ingest time using the W3C PROV Ontology~\cite{Lebo2013PROVO}: each \texttt{edm:ProvidedCHO} carries \texttt{prov:generatedAtTime}, \texttt{dcterms:hasVersion} (DDB revision id), and links to the source dataset and mapping-version agent derived from the DDB JSON metadata.
Graph-level pipeline provenance --- an Activity record per script run, capturing agent identity, version, timestamps, and an \texttt{inferenceMethod} flag (\texttt{Deterministic} / \texttt{Heuristic} / \texttt{NER}) --- is planned as future work, building on the two-layer PROV-O design developed for the companion goethe-faust corpus.

\paragraph{Update cadence.}
GEMEA tracks DDB data releases as they become available.
\todo{Confirm expected update cadence (quarterly/annual) with DDB.}

\paragraph{Persistent URI.}
The dataset is registered at \todo{[w3id.org/gemea or Zenodo DOI]}, providing a stable reference URI independent of hosting infrastructure.

\paragraph{Open source.}
The full pipeline --- rdf2jsonld, \textbf{MOCHO}, QLever configuration, SHMARQL Docker Compose, and MCP setup --- is available at \todo{[GitHub repository URL]} under an open source license.
The GEMEA index is reproducible from the DDB data dump using the provided scripts.

\paragraph{Institutional support.}
GEMEA is developed at FIZ Karlsruhe -- Leibniz Institute for Information Infrastructure in collaboration with the Karlsruhe Institute of Technology (AIFB).
The resource is hosted on institutional infrastructure.
\todo{Confirm hosting arrangement and expected availability period.}

\paragraph{Dependency management.}
The \textbf{MOCHO} alignment layer depends on \texttt{mocho.owl}, which is under active development.
Alignment rule changes upstream are tracked in the GEMEA repository and trigger a re-run of the affected named graph.
Known scope boundaries in the current alignment (Section~\ref{sec:quality}) are documented as versioned issue records.

\paragraph{v2 roadmap.}
v2 extends the corpus to the full 65M DDB objects, adds an Elasticsearch full-text index for faceted search, replaces SHMARQL with a purpose-built Next.js KG browser, and completes the GND agent and Werk linking pipeline.
The v1 QLever index and named graph structure are reused without migration.

\paragraph{Outbound URI persistence.}
The current GEMEA snapshot reflects DDB holdings as of April 2026.
Outbound links encoded as \texttt{edm:isShownAt} and \texttt{edm:object} values point to \texttt{ddb.de/item/\{id\}} URIs.
As DDB does not operate under a formal persistent URI policy, objects withdrawn or re-harvested by contributing institutions will render a proportion of these links unreachable over time.
A URI persistence layer is planned at the FIZ infrastructure level: periodic link-health checks coupled with archival fallbacks would maintain provenance navigability as upstream content evolves.
