%% 40-results.tex — Results

\section{Results}
\label{sec:results}

\subsection{Overall Pipeline Statistics}

Table~\ref{tab:stats} summarises the pipeline output on the 115{,}432-record
pilot corpus.

\begin{table}[t]
\caption{Pipeline statistics on the Goethe-\textit{Faust} pilot corpus
  (115{,}432 records, all seven DDB sectors).}
\label{tab:stats}
\centering
\begin{tabular}{lr}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Records processed                         & 115{,}432 \\
Triples emitted                           & 44{,}415{,}392 \\
Triples per record (mean)                 & 384.7 \\
\midrule
Layer 3 — dc:type W-slot matched          & 13{,}392 \,(11.6\%) \\
Layer 3 — dc:type M-slot matched          & 106 \,(0.09\%) \\
Layer 3 — fallback (mocho:Manifestation)  & 101{,}934 \,(88.3\%) \\
Records missing domain-specific type      & 52{,}274 \,(45.3\%) \\
\midrule
Layer 2 — WebResources typed (mt002)      & 103{,}496 \\
\midrule
Triples ignored (out-of-scope properties) & 3{,}592{,}475 \\
\bottomrule
\end{tabular}
\end{table}

The 88.3\% Layer~3 fallback rate reflects the real distribution of
\texttt{dc:type} values in the corpus: the majority of records either carry
no \texttt{dc:type} value, carry a value not present in the dispatch table,
or carry a value that maps only to the default Manifestation level.
This is itself an empirical finding about the corpus: most DDB records do not
carry enough type information in \texttt{dc:type} alone to resolve a
domain-specific WEMI class without sector-level heuristics or NER.

The 45.3\% rate of records missing a domain-specific type (i.e., typed as
\texttt{mocho:Manifestation} only, with no domain class from the dispatch
table or htype lookup) establishes a concrete target for downstream
enrichment.

\subsection{Field Profile and Alignment Coverage}

Table~\ref{tab:alignment-summary} summarises alignment coverage by match
method across all entity types.

\begin{table}[t]
\caption{Alignment table summary by match method across all
  (entity type, JSON key) pairs.
  \emph{in\_mocho}: RDA property present in mocho-full.owl import;
  \emph{gap}: no RDA sub-property in RDA~5.4.9 release.}
\label{tab:alignment-summary}
\centering
\begin{tabular}{lrr}
\toprule
\textbf{Match method} & \textbf{Pairs} & \textbf{In mocho} \\
\midrule
\texttt{via\_dct\_map}              & \todo{N} & \todo{N} \\
\texttt{direct}                     & \todo{N} & \todo{N} \\
\texttt{via\_dct\_map\_not\_in\_mocho} & \todo{N} & 0 \\
\texttt{unmatched}                  & \todo{N} & 0 \\
\midrule
\textbf{Total}                      & \todo{N} & \todo{N} \\
\bottomrule
\end{tabular}
\end{table}

\todo{Fill Table 2 from alignment\_ddbedm\_mocho.csv.
  Group rows by match\_method column, count rows and
  count rows where in\_mocho = True.
  Script: grep/pandas one-liner on goethe-faust/output/alignment\_ddbedm\_mocho.csv.}

\subsection{Ignored Properties}

3{,}592{,}475 triples (7.5\% of the total potential output) are explicitly
withheld because the corresponding properties fall outside the current
alignment scope.
The largest excluded groups are EDM-structural properties
(\texttt{edm:aggregatedCHO}, \texttt{edm:isShownAt},
\texttt{edm:provider} --- together 345{,}296 occurrences),
SKOS labels on Agent and Concept entities (844{,}035 occurrences),
and the DDB-specific \texttt{edm:hasMet} and \texttt{edm:hasType} properties
(316{,}806 occurrences).
A full breakdown is available in the \texttt{transform\_stats.json} output
file distributed with the pipeline.

\todo{Optionally add a small bar chart of top-10 ignored properties
  (horizontal bars, count on x-axis) — useful visual for the TPDL audience
  and fits in a half-column figure.}

\subsection{Layer~3 Dispatch Precision (Spot Checks)}

Tan et al.~\cite{tan2026mocho} report spot-check precision for two common
library-sector dispatch rules on this corpus:
\texttt{dc:type}\,=\,\textit{Buch} assigns Manifestation-level typing with
81.6\% precision ($n{=}2{,}153$);
\texttt{dc:type}\,=\,\textit{Monografie} with 99.4\% precision
($n{=}1{,}167$).
A systematic precision evaluation across all 371 \todo{confirm count}
distinct \texttt{dc:type} values in the corpus is future work.
